Referring as a collaborative process: learning to ground \\ language through language games

### 2023-08-03 paper/paper.tex
paper
Camera-ready version of Semdial paper

\begin{document}
\maketitle
\begin{abstract}
  How do artificial agents based on neural networks coordinate on a new language through referential games over 3-d scenes?
  We extended a popular CLEVR dataset to control for different combinations of features of target and distractor objects and examine the success of referential grounding learned by the agents.
  % complicated
\end{abstract}

@@

% For these experiments,
We extend the CLEVR dataset \citep{Johnson2016} by dividing the objects into one \emph{target object} and \emph{distractors}
% The target object is the main object in the scene that should be identified and communicated by the agents.
% The distractors contain similar objects to the target object.
and by controlling for the number of shared attributes between these groups as % is defined by rules researched in
in the GRE algorithm in \citep{Dale1995}.
The target object is always unique, because at least one attribute is different from the distractors.
% First, a random target object is created.
% In the next step, the distractors are created at a random location in the scene.
Each distractor can share a maximum of two attributes with the target object.
There is no restriction on the relation between distractors, hence it is possible to have multiple identical distractors in one image.
Given the ranking of features in the original GRE algorithm, the target object is therefore identifiable either by the \emph{shape} (1), the \emph{shape} and \emph{colour} (2) or the \emph{shape}, \emph{colour} and \emph{size} (3).
For each image, fixed-size bounding boxes are extracted around the centre-point of each object.
% These bounding boxes are the input for the sender and receiver in the language game.
%
% For this research, two datasets are created.
The \emph{Dale-2} dataset contains one target object and one distractor, while the \emph{Dale-5} dataset contains one target object and four distractors.
% The way, how the distractors are created is the same.
Both datasets contain 10.000 images.
Examples are shown in
Appendix \ref{sec:extended-datasets}. % shows examples of generated images with our extended code.
% Both datasets also contain ground truth information about each image, including attributes and locations for all objects.

@@

\section{Language games}
The language games were developed and run in the EGG framework \citep{Kharitonov2019}.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
Both our sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} of \citep{Lazaridou2016}, as shown in Appendix \ref*{sec:setup-language-game}.
One central difference is the production of the message.
% This research focuses on the identification of attributes for the objects and their combination, which is why
As we focus on sequences of referring expressions, made-up of different attributes, our models produce sequences of symbols for the message % (which may correspond to the attributes)
instead of a single symbol to refer to an image. % (which corresponded to different concepts of the images).
This is done by using an encoder LSTM (sender) and a decoder LSTM (receiver) to encode language descriptions.
%
% SD 2023-07-08 15:09:29 +0200: We have now already said some of this in the introduction. In the introduction we can describe the general setup of language games and here we say explicitly what is our configuration and emphasise in particular in what ways it is different from the previously implemented games. Including a list of technical details of the models is great. When restructuign sentences, copy those that fit the previous section there (and comment one out if necessary) and keep the rest here. It requires a bit of strategic thinking where to mention things in such a short paper so that we only say it once and save space.
%
%
% The target object is always the first image that is passed to the sender
% The order of the objects for the receiver is random.
Another difference is that both sender and receiver receive visual input as segmented objects rather than as two images.
The order of the objects is random, except that the first object for the sender is always the target object to be referred to.
% The figure in Appendix \ref*{sec:setup-language-game} shows the architectures of the sender and the receiver.
For the sender, the images are passed through \emph{ResNet101} \citep{He2016} and a following linear layer that reduces the dimensions to an embedding size $e_s$.
All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
This is then used as the initial state of the encoder LSTM.
After, the sequence is created through Gumbel-Softmax relaxation \citep{Jang2016}.
%
The receiver also encodes all images using \emph{ResNet101} with a following linear layer, reducing it to $e_r$.
The sequence, received by the sender is the input for its decoder LSTM, where the hidden state with a dimension of $h_r$ is randomly initialized.
After each step of the LSTM, the receiver calculates the dot product between the hidden state and all of its image encodings.
The receiver then `points' to one of the images by applying the softmax function over the results of the dot products.
The loss is calculated using the NLL-loss.
Following, the losses for all steps are summed up, and all weights of the receiver as well as the sender are updated based on this summed loss.

@@

There are five variables in the experiments that are adjusted:
(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver $h_r$ and (5) the size of the vocabulary $|V|$.
%
% The results are evaluated using accuracy if the receiver can identify the target object.
Table \ref{tab:results} shows the accuracy of the models calculated on the success of communication if the receiver can identify the target object.
A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.

@@

% The results of the experiments are summarized in Table \ref{tab:results}.
For the \emph{Dale-2} dataset it can be clearly seen that an embedding size and hidden size that are as high as the vocabulary size are beneficial for identifying the correct object.
The receiver identifies almost every sample correctly with all sizes of 10.
When the hidden and embedding sizes are increased, the guesses by the receiver are random.
Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message for the \emph{Dale-2} dataset.
%
% SD 2023-07-08 15:17:45 +0200: The hidden size is related to the vocabulary size. Both have to be optimally selected for the models to converge.
%
% The results change, when using the \emph{Date-5} dataset with four distractors.
Using \emph{Dale-5} with four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline with 23\%.
Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\% which is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
% This accuracy is still far lower than the 95\% with the \emph{Dale-2} dataset.

@@

% The second conclusion that can be drawn looks at the differences between the two datasets.
Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
Since we discriminate objects based on properties that are also distinguished in human cognition (colour, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
There are 48 possible combinations of attributes.
Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
Therefore, the models need to learn more combinations of symbols in order to attest to this higher level of detail and especially how to relate them to features in the images.
% SD 2023-07-10 18:04:24 +0200: It would be interesting to compare then the length of messages between Dale-2 and Dale-5 that becomes used.
% DK 2023-07-10: Yes, especially with a more detailed analysis of the vocabulary. Right now, the average is always 2 symbols, for both Dale-2 and Dale-5 as well as for all different vocabulary sizes. So far, I didn't check if there are sometimes longer messages and sometimes shorter messages.

@@

In our ongoing work we are investigating deeper the emerged language and the new vocabulary,
in particular whether
it uses similar categories as human language and how its words are combined to form complete messages.
In future work we will also extend the learning to the relations between entities and the features required to capture them.
% Entries for the entire Anthology, followed by custom entries
\bibliography{anthology,custom}

@@

\section{Setup of the language game}
\label{sec:setup-language-game}
%\begin{figure}
\centering
\frame{\includegraphics[width=\linewidth]{figures/architecture.png}}
% \caption{Architectures of the sender and receiver}
%  \label{fig:architectures}
% \end{figure}

>>>>

\begin{document}
\maketitle
\begin{abstract}
  How do artificial agents based on neural networks coordinate on a new language through referential games over 3-d scenes?
  We extend a popular CLEVR dataset to control for different combinations of features of target and distractor objects and examine the success of referential grounding learned by the agents.
  % complicated
\end{abstract}

@@

% For these experiments,
We extend the CLEVR dataset \citep{Johnson2016} by dividing the objects into one \emph{target object} and \emph{distractors}
% The target object is the main object in the scene that should be identified and communicated by the agents.
% The distractors contain similar objects to the target object.
and by controlling for the number of shared attributes between these groups as % is defined by rules researched in
in the GRE algorithm in \citep{Dale1995}.
The target object is always unique, because at least one attribute is different from the distractors.
% First, a random target object is created.
% In the next step, the distractors are created at a random location in the scene.
Each distractor can share a maximum of two attributes with the target object.
There is no restriction on the relation between distractors, hence it is possible to have multiple identical distractors in one image.
Given the ranking of features in the original GRE algorithm, the target object is therefore identifiable either by \emph{shape} (1), \emph{shape} and \emph{colour} (2) or \emph{shape}, \emph{colour} and \emph{size} (3).
For each image, fixed-size bounding boxes are extracted around the centre-point of each object.
% These bounding boxes are the input for the sender and receiver in the language game.
%
% For this research, two datasets are created.
The \emph{Dale-2} dataset contains one target object and one distractor, while the \emph{Dale-5} dataset contains one target object and four distractors.
% The way, how the distractors are created is the same.
Both datasets contain 10.000 images.
Examples are shown in
Appendix \ref{sec:extended-datasets}. % shows examples of generated images with our extended code.
% Both datasets also contain ground truth information about each image, including attributes and locations for all objects.

@@

\section{Language games}
The language games were developed and run in the EGG framework \citep{Kharitonov2019}.\footnote{\href{https://github.com/DominikKuenkele/MLT\_Master-Thesis}{https://github.com/DominikKuenkele/MLT\_Master-Thesis}}
Both our sender and receiver have a similar architecture to the \emph{agnostic sender} and \emph{receiver} of \citep{Lazaridou2016}, as shown in Appendix \ref*{sec:setup-language-game}.
One central difference is the production of the message.
% This research focuses on the identification of attributes for the objects and their combination, which is why
As we focus on sequences of referring expressions, made-up of different attributes, our models produce sequences of symbols for the message % (which may correspond to the attributes)
instead of a single symbol to refer to an image. % (which corresponded to different concepts of the images).
This is done by using an encoder LSTM (sender) and a decoder LSTM (receiver) to encode language descriptions.
%
% SD 2023-07-08 15:09:29 +0200: We have now already said some of this in the introduction. In the introduction we can describe the general setup of language games and here we say explicitly what is our configuration and emphasise in particular in what ways it is different from the previously implemented games. Including a list of technical details of the models is great. When restructuign sentences, copy those that fit the previous section there (and comment one out if necessary) and keep the rest here. It requires a bit of strategic thinking where to mention things in such a short paper so that we only say it once and save space.
%
%
% The target object is always the first image that is passed to the sender
% The order of the objects for the receiver is random.
Another difference is that both sender and receiver receive visual input as segmented objects rather than as two images.
The order of the objects is random, except that the first object for the sender is always the target object to be referred to.
% The figure in Appendix \ref*{sec:setup-language-game} shows the architectures of the sender and the receiver.
For the sender, the images are passed through \emph{ResNet101} \citep{He2016} and a following linear layer that reduces the dimensions to an embedding size $e_s$.
All embedded images are concatenated and passed through another linear layer to reduce the dimensions to the hidden size $h_s$.
This is then used as the initial state of the encoder LSTM.
After, the sequence of labels is generated through Gumbel-Softmax relaxation \citep{Jang2016}.
%
The receiver also encodes all images using \emph{ResNet101} with a following linear layer, reducing it to $e_r$.
The sequence, received from the sender is the input for its decoder LSTM, where a hidden state with a dimension of $h_r$ is randomly initialised.
After each step of the LSTM, the receiver calculates a dot product between the hidden state and all of its image encodings.
The receiver then `points' to one of the images by applying a softmax function over the results of the dot products.
The loss is calculated using the NLL-loss.
Following, the losses for all steps are summed up, and all weights of the receiver as well as the sender are updated based on this summed loss.

@@

There are five variables in the experiments that are adjusted:
(1) the image embedding size for the sender $e_s$, (2) the LSTM hidden size for the sender $h_s$, (3) the image/message embedding size for the receiver $e_r$, (4) the LSTM hidden size for the receiver $h_r$ and (5) the size of the vocabulary $|V|$.
%
% The results are evaluated using accuracy if the receiver can identify the target object.
Table \ref{tab:results} shows the accuracy of the models calculated on the basis of the success of communication if the receiver can identify the target object.
A random guess corresponds to 50\% in the \emph{Dale-2} dataset and 20\% in the \emph{Dale-5} dataset.

@@

% The results of the experiments are summarized in Table \ref{tab:results}.
For the \emph{Dale-2} dataset it can be clearly seen that an embedding size and hidden size that are as high as the vocabulary size are beneficial for identifying the correct object.
The receiver identifies almost every sample correctly with all sizes of 10.
When the hidden and embedding sizes are increased, the guesses by the receiver are random.
Interestingly, a vocabulary size of 10 is enough to communicate a meaningful message for the \emph{Dale-2} dataset.
%
% SD 2023-07-08 15:17:45 +0200: The hidden size is related to the vocabulary size. Both have to be optimally selected for the models to converge.
%
% The results change, when using the \emph{Date-5} dataset with four distractors.
Using \emph{Dale-5} with four distractors and with low hidden, embedding and vocabulary sizes, the agents barely pass the random baseline at 23\%.
Only increasing the vocabulary size to 100 raises the accuracy by almost 20\% to 43\% which is still considerably lower than the 95\% of the \emph{Dale-2} dataset.
% This accuracy is still far lower than the 95\% with the \emph{Dale-2} dataset.

@@

% The second conclusion that can be drawn looks at the differences between the two datasets.
Unsurprisingly, the agents have a much higher difficulty to discriminate a target object from four instead of one distractor.
Since we discriminate objects based on properties that are also distinguished in human cognition (colour, size, shape), we expect that the vocabulary onto which the agents converge reflects these categories and is therefore close to human vocabulary.
There are 48 possible combinations of attributes.
Still, for Dale-2, a vocabulary size of only 10 is enough for an almost perfect accuracy with two objects.
This hints to the fact that the agents don't describe the complete target object, but only rely on discriminative attributes between the objects.
The need for a more detailed description of discriminative attributes is higher when more distractors are involved.
Therefore, the models must learn more combinations of symbols in order to attest to this higher level of detail and how to relate them to features in the images.
% SD 2023-07-10 18:04:24 +0200: It would be interesting to compare then the length of messages between Dale-2 and Dale-5 that becomes used.
% DK 2023-07-10: Yes, especially with a more detailed analysis of the vocabulary. Right now, the average is always 2 symbols, for both Dale-2 and Dale-5 as well as for all different vocabulary sizes. So far, I didn't check if there are sometimes longer messages and sometimes shorter messages.

@@

In our ongoing work we are extending our analysis of features and agent configurations as well as we are investigating the emerged language and the new vocabulary, whether
it consists of similar categories as human language and how its words are combined to form complete messages.
In our future work we will extend the learning to relations between entities which introduce a high level referential underspecification. % and the features required to capture them.
% Entries for the entire Anthology, followed by custom entries
\bibliography{anthology,custom}

@@

\section{Setup of the language game}
\label{sec:setup-language-game}
%\begin{figure}
\centering
\frame{\includegraphics[width=\linewidth]{figures/schema.png}}
% \caption{Architectures of the sender and receiver}
%  \label{fig:architectures}
% \end{figure}
